### 2019-06-07 2-Chapters/1-Chapter/chapter1.tex
Proofread with proselint, writegood, checkwriting

    \item
    We wrote the most complete open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
    We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
    A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.

>>>>

    \item
    We wrote the most exhaustive open-source simulation library for MAB problems, called SMPyBandits, which is published online under an open-source licence \cite{SMPyBandits,SMPyBanditsJMLR}.
    We present in details its architecture and its features in Chapter~\ref{chapter:3}, along with different examples of its usage.
    A full documentation is available online, as well as exhaustive instructions to reproduce the experiments used in the rest of this thesis.

### 2019-06-07 2-Chapters/2-Chapter/chapter2.tex
Proofread with proselint, writegood, checkwriting

Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism under uncertainty and the Bayesian point-of-view.
%  underlying the approaches based on UCB, might not be necessary at all.
% With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniform perturbations.
Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$).
But instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
% https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html

@@

Many different learning algorithms have been proposed by the machine learning community,
and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
the $\kl$ function for \klUCB{} etc.
Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
%
For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms.
But when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
ideally one would like to use an algorithm that can be applied identically against any problem of such family.

>>>>

Finally, we would like to mention that a recent work \cite{KimTewari2019} presented an approach that closes the gap between the principle of optimism under uncertainty and the Bayesian point-of-view.
%  underlying the approaches based on UCB, might not be necessary at all.
% With perturbation: maybe we don't need optimism in face of uncertainty, we just need uniform sampling!
The authors of \cite{KimTewari2019} showed that their Randomized Confidence Bound (RCB) algorithm can achieve the same regret upper bounds as $\UCB_1$ or Thompson sampling, with uniform perturbations.
Their RCB algorithm ressembles $\UCB_1$, and builds the same confidence intervals $[ \mathrm{LCB}_k(t), \UCB_k(t) ]$ on arm $k$ at time $t$ (the lower confidence bound is $\mathrm{LCB}_k(t) = \widehat{\mu_k}(t) - \xi_k(t)$),
but instead of playing the arm with maximum UCB, a random sample is taken on each interval, and the played arm is the one maximizing this sample.
% https://smpybandits.github.io/docs/Policies.RandomizedIndexPolicy.html

@@

Many different learning algorithms have been proposed by the machine learning community,
and most of them depend on several parameters, for instance $\alpha$ for $\UCB_1$, the prior for Thompson sampling or BayesUCB,
the $\kl$ function for \klUCB{} etc.
Every time a new MAB algorithm $\Alg$ is introduced, it is compared and benchmarked on some bandit instance, parameterized by $\boldsymbol{\mu} = (\mu_1,\dots,\mu_K)$, usually by focusing on its expected regret $R_T^{\Alg}$.
%
For a known and specific instance, simulations help to select the best algorithm in a pool of algorithms,
but when one wants to tackle an \emph{unknown} real-world problem, one expects to be efficient against \emph{any} problem, of any size and complexity in a certain family:
ideally one would like to use an algorithm that can be applied identically against any problem of such family.

### 2019-06-07 2-Chapters/3-Chapter/SMPyBandits_paper.git/longpaper.tex
Changed \paragraph{XXX} to \paragraph{XXX.}
Proofread with proselint, writegood, checkwriting

\hypertarget{arms-and-problems}{%
\paragraph{Arms and problems}\label{arms-and-problems}}

@@

The framework is written in Python {[}@python{]}, using matplotlib
{[}@matplotlib{]} for 2D plotting, numpy {[}@numpy{]} for data storing,
random number generations and and operations on arrays, scipy
{[}@scipy{]} for statistical and special functions, and seaborn
{[}@seaborn{]} for pretty plotting and colorblind-aware colormaps.
Optional dependencies include joblib {[}@joblib{]} for parallel
simulations, numba {[}@numba{]} for automatic speed-up on some small
functions, as well as sphinx {[}@sphinx{]} for generating the
documentations. I also acknowledge the use of virtualenv
{[}@virtualenv{]} for launching simulations in isolated environments,
and jupyter {[}@jupyter{]} used with ipython {[}@ipython{]} to
experiment with the code.

>>>>

\hypertarget{arms-and-problems}{%
\paragraph{Arms and problems.}\label{arms-and-problems}}

@@

The framework is written in Python {[}@python{]}, using matplotlib
{[}@matplotlib{]} for 2D plotting, numpy {[}@numpy{]} for data storing,
random number generations and operations on arrays, scipy
{[}@scipy{]} for statistical and special functions, and seaborn
{[}@seaborn{]} for pretty plotting and colorblind-aware colormaps.
Optional dependencies include joblib {[}@joblib{]} for parallel
simulations, numba {[}@numba{]} for automatic speed-up on some small
functions, as well as sphinx {[}@sphinx{]} for generating the
documentations. I also acknowledge the use of virtualenv
{[}@virtualenv{]} for launching simulations in isolated environments,
and jupyter {[}@jupyter{]} used with ipython {[}@ipython{]} to
experiment with the code.

### 2019-06-07 2-Chapters/3-Chapter/chapter3.tex
Lots of typos changed, I'm reading all the thesis very quickly
CHAUD
FAIT CHIER
Proofread with proselint, writegood, checkwriting

\textbf{Abstract.}
%
SMPyBandits is a Python package I have developed since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%  and SMPyBandits is written in Python.
It is the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
%
This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, and a study to compare time and memory costs of the main algorithms.

@@

% \newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

@@

SMPyBandits is a Python package I have developed since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%  and SMPyBandits is written in Python.
This library is by far the most complete open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
It is hosted on GitHub, and uses the state-of-the-art development technologies, by using two online services of continuous integration to run automated tests and to build its online documentation.

@@

\paragraph{Publications.}
%
This chapter is based on our article \cite{SMPyBanditsJMLR}.
The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.

@@

    \item Algorithm 14 is
    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
    It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
    It was also proved to to obtain good ``best of both world'' performances.
    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations. Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
    \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as KL-UCB-Switch), and costs an-order-of-magnitude more memory.

@@

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb cores) & $20$ ($>$ nb cores) \\
        \hline
        $1$ repetition    & $15 \;\text{s}$ & $26 \;\text{s}$ & $43 \;\text{s}$ \\
        $10$ repetitions  & $87 \;\text{s}$ & $51 \;\text{s}$ & $76 \;\text{s}$ \\
        $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1846 \;\text{s}$ \\
        \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

>>>>

\graphicspath{{2-Chapters/3-Chapter/Images/}}
\graphicspath{{2-Chapters/3-Chapter/SMPyBandits_paper.git/plots/}}

\paragraph{Abstract.}
%
SMPyBandits is a Python package I have developed since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%  and SMPyBandits is written in Python.
It is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
%
This chapter details the organization of the library, what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparisons of the main state-of-the-art single-player MAB algorithms, and a study to compare time and memory costs of the main algorithms.

\TODOL{C'est bizarre ce problème d'espacement vertical ici... je ne sais pas comment le corriger !}

@@

@@

SMPyBandits is a Python package I have developed since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%  and SMPyBandits is written in Python.
This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
It is hosted on GitHub, and uses the state-of-the-art development technologies, by using two online services of continuous integration to run automated tests and to build its online documentation.

@@

\textbf{Publications.}
%
This chapter is based on our article \cite{SMPyBanditsJMLR}.
The code for SMPyBandits is hosted online at \texttt{\href{https://GitHub.com/SMPyBandits/SMPyBandits/}{GitHub.com/SMPyBandits/SMPyBandits}}, its documentation is at \texttt{\href{https://SMPyBandits.GitHub.io/}{SMPyBandits.GitHub.io}} or \texttt{\href{https://SMPyBandits.RTFD.io/}{SMPyBandits.RTFD.io}}, and all the library is freely publicly released under the MIT open-source License.

@@

    \item Algorithm 14 is
    \textsc{Tsallis-Inf} from \cite{Zimmert2018}, using $\alpha=1/2$ (it seems to be the most efficient of the algorithms described in the article) (\texttt{TsallisInf}).
    It is based on the online mirror descent algorithm with a Tsallis entropy regularizer, and it is anytime.
    It was also proved to obtain good ``best of both world'' performances.
    Empirically, we were unable to find any problem were it performs as well as \UCB{} (or any other efficient algorithm), and we are confident that our implementation is correct\footnote{Despite a serious amount of time spent on this implementation, maybe it has some bug which explains the poor empirical performance obtained by \textsc{Tsallis-Inf} in our simulations. Exploring in more details this kind of algorithms is very interesting, as algorithms based on online mirror descent make a rich connexion between multi-armed bandit and online convex optimization \cite{Hazan2016introduction}, and a short-term future work in SMPyBandits will be to explore different values of $\alpha$ for Tsallis entropy regularization, or other families of regularization.}.
    \textsc{Tsallis-Inf} is also slower than \UCB{} (about as slow as KL-UCB-Switch), and costs an-order-of-magnitude more memory.

@@

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Repetitions} $\;$ \textbackslash $\;$ number of jobs \texttt{N\_JOBS} & $1$ & $4$ ($=$ nb cores) & $20$ ($>$ nb cores) \\
        \hline
        $1$ repetition    & $15 \;\text{s}$ & $26 \;\text{s}$ & $43 \;\text{s}$ \\
        $10$ repetitions  & $87 \;\text{s}$ & $51 \;\text{s}$ & $76 \;\text{s}$ \\
        $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1846 \;\text{s}$ \\
        % \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

### 2019-06-07 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proofread with proselint, writegood, checkwriting

\paragraph{Logarithmic switching cost for \MCTopM.}\label{app:5:NumberSwitches}
%
Additionally to bounding the collisions in Lemma~\ref{lem:5:collisionsMCTopM},
an interesting consequence of the proof of this result is that
it also bounds the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
%
We controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
Thus, the total number of arm switches is also proven to be logarithmic, if all players uses the \MCTopM-\klUCB{} algorithm.
This additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
%
Even though minimizing the number of arms switching was not a goal,
this guarantee is interesting, in particular for Cognitive Radio applications,
where switching arms means re-configuring a radio hardware, an operation that costs energy.
An algorithm guaranteeing a small number of switches is thus interesting,
and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
%
We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17}, is an interesting future work.

>>>>

\paragraph{Logarithmic switching cost for \MCTopM.}\label{app:5:NumberSwitches}
%
Additionally to bounding the collisions in Lemma~\ref{lem:5:collisionsMCTopM},
another consequence of our proof of this result is that
it also bounds the (expected) number of \emph{switches of arms}, $SC^{\cA}(T) = \sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$.
%
We controlled the total number of transitions $(2)$, $(3)$ and $(5)$,
which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
Thus, the total number of arm switches is also proven to be logarithmic, if all players uses the \MCTopM-\klUCB{} algorithm.
This additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
%
Even though minimizing the number of arms switching was not a goal,
this guarantee is appealing, in particular for Cognitive Radio applications,
where switching arms means re-configuring a radio hardware, an operation that costs energy.
An algorithm guaranteeing a small number of switches is thus interesting,
and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
%
We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17} for single-player bandits, could also be fruitful.
